Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone

### 2019-06-16 DHKM_Response_to_Reviewers.tex

\documentclass[11pt]{article}
\usepackage{latexsym,amsfonts,amsmath,theorem,amssymb}

\newcommand{\il}{\left\langle}
\newcommand{\ir}{\right\rangle}
\newcommand{\bsb}{{\boldsymbol{b}}}
\newcommand{\bsc}{{\boldsymbol{c}}}
\newcommand{\bsgamma}{{\boldsymbol{\gamma}}}
\newcommand{\bseta}{{\boldsymbol{\eta}}}
\newcommand{\bstau}{{\boldsymbol{\tau}}}
\newcommand{\bsh}{{\boldsymbol{h}}}
\newcommand{\bsp}{{\boldsymbol{p}}}
\newcommand{\bsell}{{\boldsymbol{\ell}}}
\newcommand{\bst}{{\boldsymbol{t}}}
\newcommand{\bsr}{{\boldsymbol{r}}}
\newcommand{\bsu}{{\boldsymbol{u}}}
\newcommand{\bsv}{{\boldsymbol{v}}}
\newcommand{\bsx}{{\boldsymbol{x}}}
\newcommand{\bsq}{{\boldsymbol{q}}}
\newcommand{\bsy}{{\boldsymbol{y}}}
\newcommand{\bsz}{{\boldsymbol{z}}}
\newcommand{\bsw}{{\boldsymbol{w}}}
\newcommand{\bsS}{{\boldsymbol{S}}}
\newcommand{\bsW}{{\boldsymbol{W}}}
\newcommand{\bspitch}{{\boldsymbol{\,\pitchfork}}}
\newcommand{\bszero}{{\boldsymbol{0}}}
\newcommand{\bsone}{{\boldsymbol{1}}}
\newcommand{\rd}{\,\mathrm{d}}
\newcommand{\ri}{\mathrm{i}}
\newcommand{\re}{\mathrm{e}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\bbN}{\mathbb{N}}
\newcommand{\bbZ}{\mathbb{Z}}
\newcommand{\bbE}{\mathbb{E}}
\newcommand{\calA}{\mathcal{A}}
\newcommand{\calB}{\mathcal{B}}
\newcommand{\calC}{\mathcal{C}}
\newcommand{\calD}{\mathcal{D}}
\newcommand{\calJ}{\mathcal{J}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\calL}{\mathcal{L}}
\newcommand{\calZ}{\mathcal{Z}}
\newcommand{\calF}{\mathcal{F}}
\newcommand{\calG}{\mathcal{G}}
\newcommand{\calH}{\mathcal{H}}
\newcommand{\calI}{\mathcal{I}}
\newcommand{\calO}{\mathcal{O}}
\newcommand{\calQ}{\mathcal{Q}}
\newcommand{\calS}{\mathcal{S}}
\newcommand{\calW}{\mathcal{W}}
\newcommand{\tr}{{\tt T}}
\renewcommand{\pmod}[1]{\,(\bmod\,#1)}
\renewcommand{\mod}[1]{\,(\negthickspace\bmod #1)}
\newcommand{\ns}{\negthickspace\negthickspace}
\newcommand{\nns}{\negthickspace\negthickspace\negthickspace\negthickspace}
\newcommand{\mask}[1]{}
\newcommand{\esup}{\operatornamewithlimits{ess\,sup}}
\newcommand{\einf}{\operatornamewithlimits{ess\,inf}}
\newcommand{\e}{{\varepsilon}}
\newcommand{\setu}{{\mathfrak{u}}}
\newcommand{\setv}{{\mathfrak{v}}}
\newcommand{\setw}{{\mathfrak{w}}}
\newcommand{\setU}{{\mathfrak{U}}}
\newcommand{\sob}{{\rm Sob}}
\newcommand{\norm}[1]{\left\Vert#1\right\Vert}
\newcommand{\abs}[1]{\left\vert#1\right\vert}
\newcommand{\sset}{[s]}
\newcommand{\To}{\rightarrow}
\newcommand{\dimtr}{{\rm dim}^{\rm trnc}}
\newcommand{\dimtrint}{{\rm dim}^{\rm trnc-int}}

\begin{document}

\begin{center}

{\Large Response to the Reviewers of the Manuscript\\
``Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone''}
\end{center}

\medskip

\noindent We would like to thank you for your careful reviews of our manuscript and your helpful comments.

Following the suggestions of one of the reviewers, we have restructured the introductory parts of the paper. Thereby, we address several of the remarks 
made by both referees.

Below, we respond to the points made in your reports.

\subsection*{Referee Report 1 ``Summary. The authors study \ldots '' (2 pages):}

\begin{itemize}
\item Remarks (1)--(4): done. 
\item Remark (6): What we mean by "this example" is the example of functions defined over $[-1,1]^d$ with a Chebyshev polynomial basis. We reformulated this passage to make this point clearer. 
\item Remarks (7), (8): Due to the new structure of the paper, we hope to have addressed these issues accordingly. 
\item Remark (9): done. 
\item Remarks (11), (12): Due to the new structure of the paper, we hope to have addressed these issues accordingly. 
\item Remark (13): We put this definition into a separate paragraph. 
\item Remark (14): done. 
\item Remark (15): Due to the new structure of the paper, we hope to have addressed these issues accordingly. 
\item Remark (16): done. 
\item Remark (17): The motivation for defining this cone is stated in the beginning of the section, but we have reformulated this passage to make it clearer. 
\item Remark (18): Thanks, we have provided further details in Algorithm 4 for the step ``Compute data-driven POSD weights'', as well as for the sentence around (41). With this, the adaptive procedure is better defined, and we believe it should be called an algorithm.
\end{itemize}

\subsection*{Referee Report 2 ``The paper studies \ldots'' (4 pages):}

\begin{itemize}
\item Major issues 1.a)--1.d), 2.a), 2.b): Done. We have modified the introduction, and re-organized it into two sections. 
\item Remark (i): The ordering is not necessarily unique, as we might have weights 
that are equal. 
\item Remark (ii): the referee is correct. Thanks for pointing this out. 
\item Remark (iii): done. 
\item Remark (iv): The paper will be a chapter in a proceedings volume.
\item Remark (vii): We changed ``Choosing" to ``Identifying".
\item Remark (viii): Theoretically, we could simplify this as suggested by the reviewer; however, we would like to keep the influence of $\varepsilon$ and $R$ separate and explicit. 
\item Remark (ix): We have kept this result as a theorem, but due to the re-structuring of the paper, it is no longer part of the introduction. 
\item Remark (x): we have moved the contents of the previous Section 1.5 as suggested by the reviewer. 
\item Remark (xi): done. 
\item Remark (xix): Here, we mean a ``small number of coordinates in $f$ are important''. We have made this change in the paper.
\item Remark (xx): These guiding principles help guide the inference of weights $\boldsymbol{\lambda}$ (and its corresponding cone $\calC_{\boldsymbol{\lambda},A}$) from initial data. We have added a sentence on this on page 29. 
\item Remark (xxi): Thanks, the cone in equation (16) is now denoted as $\calC_{\boldsymbol{\lambda},A}$, and the data-inferred cone on page 30 is now denoted as $\bar{\calC}_A = \calC_{\bar{\boldsymbol{\lambda}},A}$.
\item Remark (xxii): Yes, this is correct. We have added a sentence after equation (40) clarifying this.
\item Remark (xxiii): The data-adaptive cone $\bar{\calC}$ is the earlier cone $\calC_{\boldsymbol{\lambda},A}$ in equation (16) with data-adaptive weights $\boldsymbol{\lambda} = \bar{\boldsymbol{\lambda}}$. We have clarified this in equation (42).

\end{itemize}

\bigskip

\bigskip

\noindent With kind regards,\\[0.25cm]
The authors,

\noindent Y. Ding, F.J. Hickernell, P. Kritzer, and S. Mak.

\end{document}

### 2019-06-25 DHKM_Response_to_Reviewers.tex

\subsection*{Referee Report 1 ``Summary. The authors study \ldots '' (2 pages):}

\begin{itemize}
\item Remarks (1)--(4): done. 
\item Remark (6): What we mean by "this example" is the example of functions defined over $[-1,1]^d$ with a Chebyshev polynomial basis. We reformulated this passage to make this point clearer. 
\item Remarks (7), (8): Due to the new structure of the paper, we hope to have addressed these issues accordingly. 
\item Remark (9): done. 
\item Remarks (11), (12): Due to the new structure of the paper, we hope to have addressed these issues accordingly. 
\item Remark (13): We put this definition into a separate paragraph. 
\item Remark (14): done. 
\item Remark (15): Due to the new structure of the paper, we hope to have addressed these issues accordingly. 
\item Remark (16): done. 
\item Remark (17): The motivation for defining this cone is stated in the beginning of the section, but we have reformulated this passage to make it clearer. 
\item Remark (18): Thanks, we have provided further details in Algorithm 4 for the step ``Compute data-driven POSD weights'', as well as for the sentence around (41). With this, the adaptive procedure is better defined, and we believe it should be called an algorithm.
\end{itemize}

\subsection*{Referee Report 2 ``The paper studies \ldots'' (4 pages):}

\begin{itemize}
\item Major issues 1.a)--1.d), 2.a), 2.b): Done. We have modified the introduction, and re-organized it into two sections. 
\item Remark (i): The ordering is not necessarily unique, as we might have weights 
that are equal. 
\item Remark (ii): the referee is correct. Thanks for pointing this out. 
\item Remark (iii): done. 
\item Remark (iv): The paper will be a chapter in a proceedings volume.
\item Remark (vii): We changed ``Choosing" to ``Identifying".
\item Remark (viii): Theoretically, we could simplify this as suggested by the reviewer; however, we would like to keep the influence of $\varepsilon$ and $R$ separate and explicit. 
\item Remark (ix): We have kept this result as a theorem, but due to the re-structuring of the paper, it is no longer part of the introduction. 
\item Remark (x): We have moved the contents of the previous Section 1.5 as suggested by the reviewer. 
\item Remark (xi): The cone in equation (16) is now denoted as $\calC_{\boldsymbol{\lambda},A}$. 
\item Remark (xii): We have fixed the notation in Proposition 1. Thanks for pointing this out. We have also added a Proposition 2 to discuss the relationship between our cone $\mathcal{C}_{\boldsymbol{\lambda},A}$ and a new cone $\mathcal{C}'$.
\item Remark (xiii): We have added the results with smoothness parameters $r$ in Corollary 1.
\item Remark (xiv): Thanks for pointing this out. $\kappa_i$ should be $k_i$. $k_i$ and $l_i$ are the sub-indices of $\boldsymbol{s}$ and $\boldsymbol{w}$ after ordering respectively. We could obtain the left-hand side inequality via the right-hand side special case.
\item Remark (xv): The cone in equation (28) is now denoted as $\calC_{\boldsymbol{\lambda},a,b}$. 
\item Remark (xix): Here, we mean a ``small number of coordinates in $f$ are important''. We have made this change in the paper.
\item Remark (xx): These guiding principles help guide the inference of weights $\boldsymbol{\lambda}$ (and its corresponding cone $\calC_{\boldsymbol{\lambda},A}$) from initial data. We have added a sentence on this on page 29. 
\item Remark (xxi): Thanks, the cone in equation (16) is now denoted as $\calC_{\boldsymbol{\lambda},A}$, and the data-inferred cone on page 30 is now denoted as $\bar{\calC}_A = \calC_{\bar{\boldsymbol{\lambda}},A}$.
\item Remark (xxii): Yes, this is correct. We have added a sentence after equation (40) clarifying this.
\item Remark (xxiii): The data-adaptive cone $\bar{\calC}$ is the earlier cone $\calC_{\boldsymbol{\lambda},A}$ in equation (16) with data-adaptive weights $\boldsymbol{\lambda} = \bar{\boldsymbol{\lambda}}$. We have clarified this in equation (42).

### 2019-06-25 DingHickKritzMakRICAMChapVerD.tex
updates

\DeclareMathOperator{\SOL}{SOL}
\DeclareMathOperator{\APP}{APP}
\DeclareMathOperator{\ALG}{ALG}
\DeclareMathOperator{\ERR}{ERR}
\DeclareMathOperator{\card}{card}
\newcommand{\dataN}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^n}
\newcommand{\dataNj}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^{n_j}}
\newcommand{\dataNjd}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^{n_{j^\dagger}}}
\newcommand{\ERRN}{\ERR\bigl(\dataN,n\bigr)}
\newcommand{\ERRNj}{\ERR\bigl(\dataNj,n_j\bigr)}
\newcommand{\ERRNjd}{\ERR\bigl(\dataNjd,n_{j^\dagger}\bigr)}
\DeclareMathOperator{\COST}{COST}
\DeclareMathOperator{\COMP}{COMP}
\newcommand{\hf}{\widehat{f}}
\newcommand{\hg}{\widehat{g}}
\newcommand{\tu}{\Tilde{u}}
\newcommand{\tv}{\Tilde{v}}
\newcommand{\tcalJ}{\widetilde{\calJ}}
\newcommand{\tcalK}{\widetilde{\calK}}
\newcommand{\tlambda}{\Tilde{\lambda}}
\newcommand{\tbbK}{\widetilde{\bbK}}
\newcommand{\lo}{\textup{lo}}
\newcommand{\up}{\textup{up}}
\newcommand{\inc}{\textup{in}}
\newcommand{\out}{\textup{out}}
\newcommand{\E}{\textup{e}}
\def\abs#1{\ensuremath{\left \lvert #1 \right \rvert}}
\newcommand{\normabs}[1]{\ensuremath{\lvert #1 \rvert}}
\newcommand{\bigabs}[1]{\ensuremath{\bigl \lvert #1 \bigr \rvert}}
\newcommand{\Bigabs}[1]{\ensuremath{\Bigl \lvert #1 \Bigr \rvert}}
\newcommand{\biggabs}[1]{\ensuremath{\biggl \lvert #1 \biggr \rvert}}
\newcommand{\Biggabs}[1]{\ensuremath{\Biggl \lvert #1 \Biggr \rvert}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\newcommand{\normnorm}[2][{}]{\ensuremath{\lVert #2 \rVert}_{#1}}
\newcommand{\bignorm}[2][{}]{\ensuremath{\bigl \lVert #2 \bigr \rVert}_{#1}}
\newcommand{\Bignorm}[2][{}]{\ensuremath{\Bigl \lVert #2 \Bigr \rVert}_{#1}}
\newcommand{\biggnorm}[2][{}]{\ensuremath{\biggl \lVert #2 \biggr \rVert}_{#1}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\FredNote}[1]{{\color{blue}Fred: #1}}
\newcommand{\YuhanNote}[1]{{\color{magenta}Yuhan: #1}}
\newcommand{\PeterNote}[1]{{\color{orange}Peter: #1}}
\newcommand{\SimonNote}[1]{{\color{purple}Simon: #1}}
\newcommand{\tLambda}{\widetilde{\Lambda}}
\newcommand{\bcalK}{\bar{\calK}}
\newcommand{\fapp}{f_{\text{app}}}
%\newcommand{\HickernellFJ}{Hickernell}

@@

  \author[1]{Yuhan Ding}
  \author*[2]{Fred J. Hickernell}
  \author[3]{Peter Kritzer} 
  \author[4]{Simon Mak}
  \runningauthor{Y. Ding, F. J. Hickernell, P. Kritzer, and S. Mak}
  \affil[1]{Department of Mathematics, 
  Misericorida University, 301 Lake Street, Dallas, PA 18704 USA}
  \affil[2]{Department of Applied Mathematics, Illinois Institute of Technology, RE 220, 10 W.\ 32${}\text{nd}$ Street, Chicago, IL 60616 USA}
  \affil[3]{Johann Radon Institute for Computational and Applied Mathematics (RICAM), 
  Austrian Academy of Sciences, Altenbergerstr. 69, 4040 Linz, Austria}
  \affil[4]{H. Milton Stewart School of Industrial and Systems Engineering, Georgia Institute of Technology, 755 Ferst Drive, Atlanta, GA 30332 USA}
  \title{Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone}
  \runningtitle{Adaptive Approximation}
  %\subtitle{...}
  \abstract{We study adaptive approximation algorithms for multivariate linear 
  problems where the input sets are non-convex cones. While it is known that adaptive algorithms perform 
  essentially no better than non-adaptive algorithms for convex input sets, the 
  situation may be different for non-convex sets. A typical example
  considered here is function approximation based on series expansions. Given 
  an error tolerance, we use series coefficients of the input to construct an 
  approximate solution such that the error does not exceed this tolerance. We study the situation where we can bound the norm of the input based on a pilot sample, and the situation where we keep track of the decay rate of the series coefficients of the input. Moreover, 
  we consider situations where it makes sense to infer coordinate and smoothness importance. Besides performing an error analysis, we also study the information cost of our algorithms and the computational complexity 
  of our problems, and we identify conditions under which we can avoid a curse of 
  dimensionality.}
  %\keywords{...}
  %\classification[PACS]{...}
  %\communicated{...}
  %\dedication{...}
  %\received{...}
  %\accepted{...}
  %\journalname{...}
  %\journalyear{...}
  %\journalvolume{..}
  %\journalissue{..}
  \startpage{1}
  \aop
  %\DOI{...}

@@

@@

Adaptive algorithms infer the necessary computational effort based on the function data sampled.  Adaptive algorithms may perform better than non-adaptive algorithms if the set of input functions is \emph{non-convex}. We construct adaptive algorithms for general multivariate linear problems where the input functions lie in non-convex cones.  Our algorithms use series coefficients of the input function to construct an approximate solution that satisfies an absolute error tolerance.  We show our algorithms to be essentially optimal.  We derive conditions under which the problem is tractable, i.e., the information cost of constructing the approximate solution does not increase exponentially with the dimension of the input function domain.  In the remainder of this section we define the problem and essential notation.  But first, we present a helpful example.

>>>>

\DeclareMathOperator{\SOL}{SOL}
\DeclareMathOperator{\APP}{APP}
\DeclareMathOperator{\ALG}{ALG}
\DeclareMathOperator{\ERR}{ERR}
\DeclareMathOperator{\card}{card}
\newcommand{\dataN}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^n}
\newcommand{\dataNj}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^{n_j}}
\newcommand{\dataNjd}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^{n_{j^\dagger}}}
\newcommand{\ERRN}{\ERR\bigl(\dataN,n\bigr)}
\newcommand{\ERRNj}{\ERR\bigl(\dataNj,n_j\bigr)}
\newcommand{\ERRNjd}{\ERR\bigl(\dataNjd,n_{j^\dagger}\bigr)}
\DeclareMathOperator{\COST}{COST}
\DeclareMathOperator{\COMP}{COMP}
\newcommand{\hf}{\widehat{f}}
\newcommand{\hg}{\widehat{g}}
\newcommand{\tu}{\Tilde{u}}
\newcommand{\tv}{\Tilde{v}}
\newcommand{\tcalJ}{\widetilde{\calJ}}
\newcommand{\tcalK}{\widetilde{\calK}}
\newcommand{\tlambda}{\Tilde{\lambda}}
\newcommand{\tbbK}{\widetilde{\bbK}}
\newcommand{\lo}{\textup{lo}}
\newcommand{\up}{\textup{up}}
\newcommand{\inc}{\textup{in}}
\newcommand{\out}{\textup{out}}
\newcommand{\E}{\textup{e}}
\def\abs#1{\ensuremath{\left \lvert #1 \right \rvert}}
\newcommand{\normabs}[1]{\ensuremath{\lvert #1 \rvert}}
\newcommand{\bigabs}[1]{\ensuremath{\bigl \lvert #1 \bigr \rvert}}
\newcommand{\Bigabs}[1]{\ensuremath{\Bigl \lvert #1 \Bigr \rvert}}
\newcommand{\biggabs}[1]{\ensuremath{\biggl \lvert #1 \biggr \rvert}}
\newcommand{\Biggabs}[1]{\ensuremath{\Biggl \lvert #1 \Biggr \rvert}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\newcommand{\normnorm}[2][{}]{\ensuremath{\lVert #2 \rVert}_{#1}}
\newcommand{\bignorm}[2][{}]{\ensuremath{\bigl \lVert #2 \bigr \rVert}_{#1}}
\newcommand{\Bignorm}[2][{}]{\ensuremath{\Bigl \lVert #2 \Bigr \rVert}_{#1}}
\newcommand{\biggnorm}[2][{}]{\ensuremath{\biggl \lVert #2 \biggr \rVert}_{#1}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\FredNote}[1]{{\color{blue}Fred: #1}}
\newcommand{\YuhanNote}[1]{{\color{magenta}Yuhan: #1}}
\newcommand{\PeterNote}[1]{{\color{orange}Peter: #1}}
\newcommand{\SimonNote}[1]{{\color{purple}Simon: #1}}
\newcommand{\DHKMchange}[1]{{\color{red}{#1}}}

\newcommand{\tLambda}{\widetilde{\Lambda}}
\newcommand{\bcalK}{\bar{\calK}}
\newcommand{\fapp}{f_{\text{app}}}
%\newcommand{\HickernellFJ}{Hickernell}

@@

  \author[1]{Yuhan Ding}
  \author*[2]{Fred J. Hickernell}
  \author[3]{Peter Kritzer} 
  \author[4]{Simon Mak}
  \runningauthor{Y. Ding, F. J. Hickernell, P. Kritzer, and S. Mak}
  \affil[1]{Department of Mathematics, 
  Misericorida University, 301 Lake Street, Dallas, PA 18704 USA}
  \affil[2]{Department of Applied Mathematics, Illinois Institute of Technology, RE 220, 10 W.\ 32${}\text{nd}$ Street, Chicago, IL 60616 USA}
  \affil[3]{Johann Radon Institute for Computational and Applied Mathematics (RICAM), 
  Austrian Academy of Sciences, Altenbergerstr. 69, 4040 Linz, Austria}
  \affil[4]{H. Milton Stewart School of Industrial and Systems Engineering, Georgia Institute of Technology, 755 Ferst Drive, Atlanta, GA 30332 USA}
  \title{Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone}
  \runningtitle{Adaptive Approximation}
  %\subtitle{...}
  \abstract{We study adaptive approximation algorithms for general multivariate linear 
  problems where the sets of input functions are non-convex cones. While it is known that adaptive algorithms perform 
  essentially no better than non-adaptive algorithms for \DHKMchange{convex and symmetric} input sets, the 
  situation may be different for non-convex sets. A typical example
  considered here is function approximation based on series expansions. Given 
  an error tolerance, we use series coefficients of the input to construct an 
  approximate solution such that the error does not exceed this tolerance. We study the situation where we can bound the norm of the input based on a pilot sample, and the situation where we keep track of the decay rate of the series coefficients of the input. Moreover, 
  we consider situations where it makes sense to infer coordinate and smoothness importance. Besides performing an error analysis, we also study the information cost of our algorithms and the computational complexity 
  of our problems, and we identify conditions under which we can avoid a curse of 
  dimensionality.}
  %\keywords{...}
  %\classification[PACS]{...}
  %\communicated{...}
  %\dedication{...}
  %\received{...}
  %\accepted{...}
  %\journalname{...}
  %\journalyear{...}
  %\journalvolume{..}
  %\journalissue{..}
  \startpage{1}
  \aop
  %\DOI{...}

@@

\DHKMchange{

@@

Adaptive algorithms infer the necessary computational effort based on the function data sampled.  Adaptive algorithms may perform better than non-adaptive algorithms if the set of input functions is \emph{non-convex} \DHKMchange{or \emph{non-symmetric}}. \DHKMchange{Here we focus on the former situation and} construct adaptive algorithms for general multivariate linear problems where the input functions lie in non-convex cones. \DHKMchange{By a cone, we mean a set of functions satisfying that for each element also all its multiples lie in the set. We will study different variants of cone here, the precise definitions will be given below.}

Our algorithms use a finite number of series coefficients of the input function to construct an approximate solution that satisfies an absolute error tolerance.  We show our algorithms to be essentially optimal.  We derive conditions under which the problem is tractable, i.e., the information cost of constructing the approximate solution does not increase exponentially with the dimension of the input function domain.  In the remainder of this section we outline the general idea of the paper.

We study the approximation of a solution operator $\SOL:\calF \to \calG$, which maps $f\in\calF$ to some $g\in\calG$, where we assume that $\calF$ and and $\calG$ are Banach spaces with norms $\norm[\calF]{\cdot}$ and $\norm[\calG]{\cdot}$, respectively. Our goal is to define a suitable set of input functions $\calC$, and an algorithm $\ALG(f,\varepsilon)$ for $f\in\calC$ that satisfies the error criterion
\begin{equation} \label{DHKM:err_crit}
    \norm[\calG]{\SOL(f) - \ALG(f,\varepsilon)} \le \varepsilon \qquad \forall \varepsilon > 0, \ f \in \calC,
\end{equation}
where  $\varepsilon$ denotes the error tolerance. Such an algorithm will  contain a rule for choosing $n$---depending on $f$ and $\varepsilon$.  The objectives of this chapter are to 
\begin{itemize}
    \item construct such a rule, 
    \item choose a set $\calC$ of input functions for which the rule is valid,  
    \item characterize the information cost of $\ALG$, 
    \item determine whether $\ALG$ has optimal information cost, and 
    \item understand the dependence of this cost on the number of input variables, $d$, as well as the error tolerance, $\varepsilon$.
\end{itemize}

We assume in this paper that the functions $f\in\calF$ can be represented in terms of a basis representation as $f = \sum_{\bsk \in \mathbb{K}} \hf(\bsk) u_{\bsk}$, and we allow algorithms $\ALG (f,\varepsilon)$ to use $n=n(f,\varepsilon)$ coefficients $\hf (\bsk)$ of this representation. I.e., 
we will study algorithms which have an output of the form 
\begin{equation*} 
  \sum_{i=1}^{n(f,\varepsilon)}\hf(\bsk_i) u_{\bsk_i} \qquad \forall f \in \calF, \ n \in \N,
\end{equation*}
where an oracle provides $\hf(\bsk_1),\ldots,\hf (\bsk_{n(f,\varepsilon)})$ 
with $n(f,\varepsilon)$ being chosen based on a stopping rule. 
I.e., we assume that it is possible to sample the series coefficients of the input function.  This is a \DHKMchange{more} restrictive assumption than being able to sample any linear functional, but it is \DHKMchange{less} restrictive than only being able to sample function values.  An important future problem is to extend the theory in this chapter to the case where the only function data available are function values.

We call an algorithm \emph{non-adaptive} if it is such that $n(f,\varepsilon)=n(\varepsilon)$, independently of the particular choice of $f\in\calC$.

The information cost of $\ALG(f,\varepsilon)$ is denoted $\COST(\ALG,f,\varepsilon)$ and defined as the number of function data---in our situation, series coefficients---required by $\ALG(f,\varepsilon)$.  For adaptive algorithms this cost varies with the input function $f$.

In the following, let $\calB_{R} : = \{ f \in \calF : \norm[\calF]{f} \le R \}$ denote the ball of radius $R$ in the space of input functions.

We now define the information cost of the algorithm in general, recognizing that it will tend to depend on $\norm[\calF]{f}$:
\begin{equation*}
    \COST(\ALG, \calC, \varepsilon,R) : = \max_{f \in \calC \cap \calB_{R}} \COST(\ALG,f,\varepsilon).
\end{equation*}
Note that while the cost depends on $\norm[\calF]{f}$, $\ALG(f,\varepsilon)$ has no knowledge of $f$ beyond the fact that it lies in $\calC$.  It is common for $\COST(\ALG, \calC, \varepsilon,R)$ to be $\calO(\varepsilon^{-p})$, or perhaps asymptotically $c\log(1 + \varepsilon^{-1})$.

\medskip

\DHKMchange{A crucial concept is the set of all possible algorithms that may be constructed using series coefficients and that \emph{satisfy error criterion \eqref{DHKM:err_crit}}. We denote this set by $\calA(\calC)$. }

Furthermore, we define the \emph{computational complexity} of a problem as the information cost of the best algorithm:
\begin{equation*}
    \COMP(\calA(\calC), \varepsilon,R) := \min_{\ALG \in \calA(\calC)} \COST(\ALG, \calC, \varepsilon,R) .
\end{equation*}
These definitions follow the information-based complexity literature \cite{TraWer98, TraWasWoz88}.
We define an algorithm to be \emph{essentially optimal} if there exist some fixed positive $\omega$, $\varepsilon_{\max}$, and $R_{\min}$ for which
\begin{multline} \label{DHKM:EssentialOpt}
    \COST(\ALG, \calC, \varepsilon,R) \le \COMP(\calA(\calC), \omega \varepsilon,R) \\ \forall \, \varepsilon \in (0, \varepsilon_{\max}], \ R \in [R_{\min}, \infty).
\end{multline}
If the complexity of the problem is $\calO(\varepsilon^{-p})$, the cost of an essentially optimal algorithm is also $\calO(\varepsilon^{-p})$. If the complexity of the problem is asymptotically $c \log(1 + \varepsilon^{-1})$, then  the cost of an essentially optimal algorithm is also asymptotically $c \log(1 + \varepsilon^{-1})$. 
We will show that our adaptive algorithms presented in Sections \ref{DHKM:pilot_sec} and \ref{DHKM:tracking_sec}  are essentially optimal.

We will demonstrate below that non-adaptive algorithms are optimal when the set of inputs is chosen to be $\calC = \calB_R$. However, typically one cannot bound the norm of the input function a priori, so non-adaptive algorithms may be impractical. 
  This will lead to \emph{adaptive} algorithms $\ALG$ satisfying error criterion \eqref{DHKM:err_crit}.  For such algorithms, the set of allowable input functions, $\calC$, will be a \emph{cone}, not a ball. By a cone, we mean a set of functions satisfying that for each element also all its multiples lie in the set. We will study different variants of cones in different sections of this paper. The first definition of a cone will be given in \eqref{DHKM:pilot_cone}, 
where the idea is that the consists of $f\in\calF$ such that 
$$
  \norm[\calF]{f}\le h (\hf(\bsk_1),\ldots,\hf (\bsk_{n_1}))
$$
for a suitably chosen homogeneous function $h:\bbR\to [0,\infty)$. This means that the norm of $f$ can be bounded well in terms of 
a \emph{pilot sample} of size $n_1\in\bbN$, see \eqref{DHKM:pilot_cone} for the precise definition. In Section \ref{DHKM:SecAdapAlgTrackDecay}, we will modify the definition of cones such that we can keep track of the decay rate of the series coefficients of the input function (see \eqref{DHKM:TrackConeDef}), 
and in Section \ref{DHKM:smoothimportance_sec} we allow a situation 
where we infer coordinate and smoothness importance from some initial data, thus making the definition of the cone more practical (see  
\eqref{DHKM:pilot_cone2}).

Note that algorithms satisfying error criterion \eqref{DHKM:err_crit} cannot exist for $\calC = \calF$. Any algorithm must require a finite sample size, even if it is huge.  Then, there must exist some $f \in \calF$ that looks exactly like the zero function to the algorithm but for which $\norm[\calG]{\SOL(f)}$ is arbitrarily large.  Thus, algorithms satisfying the error criterion  exist only for some strict subset of $\calF$.  Identifying that subset is both an art and a science.

The rest of the paper is structured as follows. In the next section, we will formulate our problem more formally, and we point out some results for non-adaptive algorithms with inputs from balls. Furthermore, we illustrate our theory by 
a particular example, and mention some definitions regarding \emph{tractability}.

In Section \ref{DHKM:pilot_sec}, we define a cone of input functions, $\calC$, in \eqref{DHKM:pilot_cone} whose norms can be bounded above in terms of the series coefficients obtained from a pilot sample.  Adaptive Algorithm \ref{DHKM:PilotConeAlg} is shown to be optimal for this $\calC$.  We also identify necessary and sufficient conditions for tractability.

Section \ref{DHKM:tracking_sec} considers the situation where function data is relatively inexpensive, and we track the decay rate of the series coefficients.  Adaptive Algorithm \ref{DHKM:TrackConeAlg} is shown to be optimal in this situation.

Section \ref{DHKM:smoothimportance_sec} considers the case where the most suitable weights $\bslambda$ are not known a priori and are instead inferred from function data.  Adaptive Algorithm \ref{DHKM:InfPilotConeAlg} combines this inference step with Algorithm \ref{DHKM:PilotConeAlg} to construct an approximation that satisfies the error criterion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Problem Formulation and an Illustrative Example}\label{DHKM:secprobdef}

\PeterNote{Can we find a better heading for this section?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{General problem formulation}

Now, we define our problem more formally.  The solution operator $\SOL$ maps the input function to an output, $\SOL:\calF \to \calG$.  The Banach spaces of inputs and outputs are defined by series expansions:
\DHKMchange{
\begin{gather*}
    \calF := \left \{f = \sum_{\bsk \in \mathbb{K}} \hf(\bsk) u_{\bsk} : \norm[\calF]{f} : = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad\mbox{with}\quad 1 \le \rho \le \infty, 
\end{gather*}
and
\begin{gather*}
    \calG := \left \{g = \sum_{\bsk \in \mathbb{K}} \hg(\bsk) v_{\bsk} : \norm[\calG]{g} : = \norm[\tau]{\bigl(  \hg(\bsk)  \bigr)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad\mbox{with}\quad 1 \le \tau \le \rho.
\end{gather*}
}
Here, $\{u_{\bsk}\}_{\bsk \in \mathbb{K}}$ is a basis for the input Banach space $\calF$, $\{v_{\bsk}\}_{\bsk \in \mathbb{K}}$ is a basis for the output Banach space $\calG$, $\mathbb{K}$ is a countable index set, and $\bslambda = (\lambda_\bsk)_{\bsk \in \mathbb{K}}$ is the sequence of weights. These bases are defined to match the solution operator:
\begin{equation} \label{DHKM:basis_relate}
    \SOL(u_{\bsk}) = v_{\bsk} \qquad \forall \bsk \in \mathbb{K}.
\end{equation}
The $\lambda_{\bsk}$ represent the importance of the series coefficients of the input function.  The larger $\lambda_{\bsk}$ is, the more important $\hf(\bsk)$ is.

To facilitate our derivations below, we establish the following lemma via H\"older's inequality:
